---
title: "BBRKC GMACS to RTMB Porting Notes"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
execute:
  echo: true
  warning: false
  message: false
---

## Purpose

This document tracks the staged conversion of the BBRKC GMACS ADMB model to an
RTMB implementation. The current goal is not to translate all of `gmacs.tpl` at
once. The goal is to create a reproducible sequence of parity checks where each
deterministic model block is ported and compared with existing ADMB output
before the corresponding likelihood components are activated.

The BBRKC case is read from `examples/BBRKC` and currently uses:

- `gmacs.dat` for file routing and run controls.
- `bbrkc_21_1.dat` for dimensions, catch, survey, composition, growth, and
  environmental data.
- `bbrkc_21_1.ctl` for phases, selectivity structure, mortality controls, and
  likelihood weights.
- `gmacs.pin` for the fitted full parameter state.
- `gmacs.rep` and `Gmacsall.out` as ADMB parity references.

## Current RTMB Files

The current scaffold lives at the repository root.

| File | Role |
|---|---|
| `R/read_gmacs_bbrkc.R` | Reads BBRKC GMACS inputs, fitted `.pin` values, ADMB likelihood blocks, selectivity reports, mortality reports, growth reports, and numbers-at-size reports. |
| `R/gmacs_rtmb_nll.R` | Builds RTMB data and parameter lists, maps full `.pin` values to active parameters, and contains deterministic model blocks. |
| `run_bbrkc.R` | Runs parser validation, constructs the mapped RTMB object, and prints deterministic parity diagnostics. |
| `optimize_bbrkc.R` | Runs a short `nlminb()` optimization smoke test from the mapped 366-parameter RTMB object. |

## Parameter Mapping

The full BBRKC `.pin` file contains 560 numeric entries. ADMB reports 366 active
estimated parameters. The function `make_bbrkc_rtmb_map()` builds an RTMB
`map` from the BBRKC control phases and maps inactive entries to `NA`.

Special cases:

- `log_fdev` and `log_fdov` activity follows the corresponding fleet-level
  `log_fbar` and `log_foff` phases.
- `log_vn` uses the GMACS composition aggregator, reducing 13 input composition
  controls to 8 effective sample-size entries.
- `logit_rec_prop_est` is treated like an ADMB dev vector, so one element is
  mapped off to reproduce 45 active entries from 46 values.
- BBRKC uses `FREEPARSSCALED` initial numbers. The first scaled `logN0` entry
  is fixed at zero and `theta[13:91]` supplies the remaining 79 sex/shell/size
  offsets.

Expected check:

```{r}
#| eval: false
source("R/read_gmacs_bbrkc.R")
source("R/gmacs_rtmb_nll.R")

inputs <- read_bbrkc_inputs("examples/BBRKC")
map <- make_bbrkc_rtmb_map(inputs)
count_active_map_parameters(map)
```

Expected result: `366`.

## Deterministic Status

The current RTMB scaffold ports BBRKC selectivity, fishing mortality, natural
mortality, total mortality/survival setup, growth, recruitment, initial numbers,
the annual/seasonal population update, catch predictions, and catch-at-size
predictions. Catch, robust size-composition, and recruitment likelihoods are
now active in the RTMB objective and checked against the ADMB `nloglike`
reference. Survey/index predictions and likelihoods are now active and matched
against the ADMB `#Index_fit_summary` and raw `nloglike` reference.

| Block | RTMB function | ADMB reference | Status |
|---|---|---|---|
| Parameter map | `make_bbrkc_rtmb_map()` | ADMB active parameter count | 560 `.pin` entries mapped to 366 active parameters |
| Capture selectivity | `calc_bbrkc_capture_selectivity()` | `gmacs.rep`, `slx_capture` | Matches report rounding |
| Retention and discard selectivity | `calc_bbrkc_retention_selectivity()` | `gmacs.rep`, `slx_retaind`, `slx_discard` | Matches report rounding |
| Fishing mortality | `calc_bbrkc_fishing_mortality()` | `Gmacsall.out`, fully selected F and continuous F-at-size | Matches ADMB numerical output |
| Natural mortality | `calc_bbrkc_natural_mortality()` | `Gmacsall.out`, natural mortality by class | Matches ADMB numerical output |
| Total mortality | `calc_bbrkc_total_mortality()` | `Gmacsall.out`, continuous and discrete total mortality by class | Matches ADMB numerical output |
| Molt probability | `calc_bbrkc_molting_probability()` | `Gmacsall.out`, molt probability | Matches ADMB numerical output |
| Growth transition | `calc_bbrkc_growth_transition()` | `Gmacsall.out`, growth transition matrix | Matches ADMB numerical output and uses `RTMB::pgamma()` so active growth parameters affect the tape |
| Recruitment | `calc_bbrkc_recruits()` | `Gmacsall.out`, overall summary | Matches ADMB numerical output to printed precision |
| Initial numbers | `calc_bbrkc_initial_numbers()` | `Gmacsall.out`, `N(...)` at first output season | FREEPARSSCALED normalized logN offsets ported |
| Population update | `calc_bbrkc_population_numbers()` | `Gmacsall.out`, `N(total)`, sex, and shell blocks | Matches ADMB numerical output to printed precision |
| Catch observations | `calc_bbrkc_predicted_catch()` | `Gmacsall.out`, `Catch_fit_summary` | Matches ADMB numerical output to printed precision; AD-safe flat vector is taped and reported |
| Catch-at-size | `calc_bbrkc_predicted_size_compositions()` | `Gmacsall.out`, `Size_fit_summary` | Matches ADMB numerical output to printed precision; AD-safe flat vector is taped and reported |
| Catch likelihood | `catch_likelihood()` inside `gmacs_rtmb_nll_factory()` | `Gmacsall.out`, `nloglike` row 1 | Matches ADMB raw likelihood components |
| Survey index | `calc_bbrkc_predicted_survey()` | `Gmacsall.out`, `#Index_fit_summary` | Matches ADMB predicted index output to printed precision |
| Survey likelihood | `index_likelihood()` inside `gmacs_rtmb_nll_factory()` | `Gmacsall.out`, `nloglike` row 2 | Matches ADMB raw likelihood components |
| Length likelihood | `length_likelihood()` inside `gmacs_rtmb_nll_factory()` | `Gmacsall.out`, `nloglike` row 3 | Robust multinomial likelihood matches ADMB raw components |
| Recruitment likelihood | `recruitment_likelihood()` inside `gmacs_rtmb_nll_factory()` | `Gmacsall.out`, `nloglike` row 4 | Uses ADMB-style `res_recruit`; matches ADMB raw components |
| Penalties | `calc_bbrkc_penalties()` | `Gmacsall.out`, `nlogPenalty` | Raw penalty vector matches ADMB; weighted vector enters the RTMB objective |
| Priors | `calc_bbrkc_prior_density()` | `Gmacsall.out`, `priorDensity` | Prior-density vector matches ADMB and enters the RTMB objective |

## Selectivity Details

Capture selectivity:

- All BBRKC capture components are logistic.
- Logistic curves are normalized to the terminal size class because the control
  file sets `slx_max_at_1 = 1`.
- NMFS trawl selectivity is embedded within BSFRF selectivity, matching
  `slx_incl = 6`.

Retention selectivity:

- Pot fishery male retention is logistic.
- The Pot male retention curve changes parameter block in 2005.
- Pot female and all non-Pot retention components are effectively zero
  retention for BBRKC.
- Discard selectivity is computed as `log(1 - retained + 1e-08)`, matching the
  ADMB template.
- The 1975 Pot male high-grading adjustment from `Asymret` is applied.

Current parity command:

```{bash}
#| eval: false
Rscript run_bbrkc.R
```

Current expected selectivity differences are on the order of GMACS report
rounding:

| Component | Maximum absolute difference |
|---|---:|
| Capture selectivity | about `5e-05` |
| Retained selectivity | about `4e-05` |
| Discard selectivity | about `4e-05` |

## Mortality Details

Fishing mortality:

- `log_fbar` supplies the fleet-level mean fishing mortality.
- `log_fdev` supplies male annual/seasonal deviations for each fleet.
- Female F uses the fleet-level `log_foff` offset and the active `log_fdov`
  deviations where female catches occur.
- Vulnerability is computed from capture selectivity, retention selectivity,
  and discard mortality.
- Both fully selected F and continuous F-at-size are compared with
  `Gmacsall.out`.

Natural mortality:

- BBRKC uses GMACS natural mortality type 6.
- Male base M is `theta[1]`.
- Female base M is relative to male M through `theta[2]`.
- The control file defines two natural-mortality nodes per sex, with the first
  deviation active over 1980-1984.
- `Mdev_spec = -1` shares the first male M-deviation parameter with the female
  block, matching the ADMB template.
- No BBRKC size-specific M multiplier is active.

Total mortality:

- Continuous total mortality is `Z = m_prop * M + F`.
- Discrete total mortality is `Z2 = m_prop * M + F2`.
- The survival diagonal `S` is computed from `Z` for the BBRKC season type.

## Growth Details

Growth parameter unpacking:

- BBRKC uses growth transition option 3, where growth increments are gamma
  distributed.
- The growth increment model is empirical by size class.
- Male growth has one transition block.
- Female growth has three transition blocks, with changes in 1983 and 1994.
- `Gscale` is handled with the BBRKC relative-beta setting, so female periods 2
  and 3 inherit the period-1 scale when their fitted relative values are zero.

Molt probability:

- BBRKC uses logistic molt probability.
- Both sexes define two molt-probability blocks, with a change in 1980.
- Female fitted values make molt probability effectively one for all modeled
  sizes and years.

Growth transition:

- The transition matrix is computed from active growth parameters using the
  GMACS gamma-increment formula.
- GMACS prints `trans(growth_transition)`, so the ADMB parser explicitly treats
  printed rows as destination size classes and printed columns as source size
  classes.
- Gamma CDF calls use `RTMB::pgamma()` through the local `ad_pgamma()` wrapper,
  avoiding the previous numeric precompute from `gmacs.pin`.

## Recruitment and Population Details

Recruitment:

- Annual recruitment is split by sex with `logit_rec_prop_est`, matching the
  ADMB dev-vector indexing.
- The first-year initial recruits use `logRini` for `FREEPARSSCALED` initial
  numbers.
- Recruitment size distribution is computed from the GMACS gamma recruitment
  formula and truncated at the sex-specific maximum recruitment size.
- Recruitment size probabilities also use `RTMB::pgamma()` through
  `ad_pgamma()`, so the distribution responds to active `theta` recruitment
  size parameters on the tape.

Initial numbers:

- The group order follows the ADMB loops over sex, maturity, and shell.
- For BBRKC this produces four groups: male new shell, male old shell, female
  new shell, and female old shell.
- The first `logN0` offset is fixed at zero; the remaining 79 offsets are read
  from `theta[13:91]`.
- Initial numbers are normalized as
  `exp(logRini + logN0) / sum(exp(logN0))`.

Population update:

- BBRKC uses the GMACS Case-D branch: two shell conditions, one maturity state,
  and no terminal molt.
- Survival is applied before growth and recruitment.
- In the growth season, molting crabs transition through the sex- and
  period-specific growth matrix; new-shell nonmolters become old-shell crabs.
- In the recruitment season, recruits enter the new-shell group by sex and
  recruitment size distribution.
- Output comparisons use `season_N` from the data file and compare total,
  sex-specific, and shell-specific numbers-at-size.

## Catch Prediction Details

Mean weight:

- BBRKC uses length-weight input type 2: fixed mean weight-at-size vectors by
  sex and maturity.
- The RTMB parser reads mean weight, maturity, and legal vectors from the
  control file because catch observations are in mass units.

Catch observations:

- `calc_bbrkc_predicted_catch()` ports the Baranov prediction in
  `calc_predicted_catch()`.
- Retained catch uses capture selectivity times retention.
- Discarded catch uses capture selectivity times one minus retention.
- Total catch uses capture selectivity alone.
- Sex-specific and combined-sex catch rows are both handled.
- Catch rows with zero observed catch and positive effort use the GMACS
  catchability calculation for `obs_catch_effort`.

Catch-at-size:

- `calc_bbrkc_predicted_size_compositions()` ports the catch-like branch of
  `calc_predicted_composition()`.
- The function predicts each of the 13 input composition series, then applies
  the GMACS composition appender to produce the 8 modified series used in the
  likelihood and report.
- Series with `lf_catch_in == 2` are survey-like composition predictions and
  use numbers at size without the catch mortality multiplier.
- The 2021 season-1 survey-like size-composition rows use the propagated
  `nyr + 1` population state and the terminal modeled-year selectivity.

Taping status:

- Catch and catch-at-size prediction helpers now use AD-compatible vector
  construction.
- `MakeADFun()` tapes these predictions and reports `model$pre_catch`,
  `model$res_catch`, `model$log_q_catch`, and `model$pre_size_comps`.
- The composition report is flattened and accompanied by `model$size_comp_rows`
  and `model$size_comp_cols`, which reconstruct the 8 modified composition
  series.

## Next Porting Steps

### 1. Prediction Blocks

Port or refine the remaining deterministic and objective blocks in this order:

1. Survey-like size-composition branch review for `lf_catch_in == 2`, using
   the now-active survey prediction logic as a guide.
2. Optional growth likelihood scaffolding if active growth observations are
   introduced.
3. Broader validation against additional GMACS model configurations.

These should continue to be checked in small pieces against `Gmacsall.out` and
`gmacs.rep` before likelihood components are activated.

### 2. Predictions and Likelihoods

The active likelihood blocks are:

- catch likelihood
- survey index likelihood
- robust length-composition likelihood
- recruitment likelihood

Remaining likelihood work:

- survey-like size-composition likelihood checks after survey composition
  predictions are reconciled
- growth likelihood only if active growth observations are introduced

`run_bbrkc.R` now performs this audit directly after taping the objective. Each
deterministic prediction change should be followed by a run that compares the
active `nloglike` components, raw `nlogPenalty`, and `priorDensity` vectors
against the ADMB blocks read from `Gmacsall.out`.

## Optimization Smoke Test

The short optimizer smoke test uses the same BBRKC parser, mapped parameter
list, RTMB objective, and 366-parameter map. It is intentionally capped at a
small number of iterations so it checks that the taped objective and gradient
are usable without treating the result as a production fit.

```{bash}
#| eval: false
Rscript optimize_bbrkc.R
```

Current smoke-test output:

```text
active parameter length: 366
initial objective: -25377.33
finite gradient: TRUE
max abs gradient: 28.4894
optimizer convergence: 1
optimizer message: iteration limit reached without convergence (10)
final objective: -25393.89
objective change: -16.55864
```

## Run Log

The current successful run reports:

```text
mapped active parameters: 366
RTMB object taped
active parameter length: 366
initial objective: -25377.33
likelihood components:
  catch RTMB=-271.2013 ADMB=-271.2013 diff=1.977e-07
  index RTMB=-37.85266 ADMB=-37.85266 diff=5.767e-07
  length RTMB=-25524.42 ADMB=-25524.42 diff=-3.529e-07
  recruitment RTMB=145.2397 ADMB=145.2397 diff=-6.589e-08
  growth RTMB=0 ADMB=0 diff=0
raw penalty max abs diff: 1.023e-06
raw penalty sum: 4995.897
weighted penalty sum: 43.9388
prior density max abs diff: 1.548e-07
prior density sum: 266.9703
capture selectivity max abs diff: 4.972e-05
retained selectivity max abs diff: 4.164e-05
discard selectivity max abs diff: 4.163e-05
fully selected F max abs diff: 3.347e-08
F at size max abs diff: 2.037e-08
natural mortality max abs diff: 4.575e-09
continuous total mortality max abs diff: 2.037e-08
discrete total mortality max abs diff: 4.235e-08
molt probability max abs diff: 6.996e-09
growth transition max abs diff: 8.162e-09
recruits max abs diff: 1.947
N(total) max abs diff: 1.784
N(males) max abs diff: 0.438
N(females) max abs diff: 1.742
N(males_new) max abs diff: 0.438
N(females_new) max abs diff: 1.742
N(males_old) max abs diff: 0.1004
N(females_old) max abs diff: 2.993e-37
catch predicted max abs diff: 7.484e-05
catch residual max abs diff: 1.145e-08
index predicted max abs diff: 0.0006836
index q max abs diff: 1.483e-05
size composition predicted max abs diff: 4.999e-05
```

The current RTMB objective includes catch, survey index, length-composition, and
recruitment likelihoods plus the weighted ADMB penalty contribution. Raw
likelihood components match the ADMB `nloglike` block, and raw penalty
components match the ADMB `nlogPenalty` block to printed precision. The
`priorDensity` vector is active and matches ADMB to numerical precision. The
abundance differences above are on the scale of individual crabs while the
reported arrays contain millions, so they are consistent with ADMB text-output
precision.

## Appendix: Model Equations

This appendix records the equation set represented by the current RTMB scaffold
and the remaining GMACS equations that still need to be activated. The
conversion is not complete, but the BBRKC population dynamics, growth
transition matrices, recruitment size distributions, catch predictions,
catch-at-size predictions, catch likelihood, robust length-composition
likelihood, survey index likelihood, and recruitment likelihood are ported and
taped; weighted penalties and prior densities are active.

### A. Notation

Indices:

- $i$ indexes model years.
- $j$ indexes seasons.
- $h$ indexes sex.
- $m$ indexes maturity state.
- $o$ indexes shell state.
- $g = (h,m,o)$ indexes the combined GMACS population group.
- $\ell$ indexes size classes.
- $k$ indexes fisheries or survey fleets.
- $r$ indexes rows in the observation tables.
- $a$ indexes modified size-composition series after GMACS aggregation.

Main state and prediction quantities:

- $N_{g,i,j,\ell}$ is abundance at size at the start of year $i$ and season
  $j$.
- $S_{k,h,i,\ell}$ is capture selectivity.
- $R_{k,h,i,\ell}$ is retention probability.
- $V_{k,h,i,\ell}$ is the fully selected fishing mortality multiplier at size.
- $F_{h,i,j,\ell}$ is fishing mortality at size after summing over active
  fleets.
- $M_{h,m,i,\ell}$ is natural mortality at size.
- $Z_{h,m,i,j,\ell}$ is total mortality for continuous seasons.
- $G_{h,i,\ell,\ell'}$ is the growth transition matrix.
- $P^{\mathrm{molt}}_{h,i,\ell}$ is molt probability.
- $\widehat{C}_r$ is predicted catch for observation row $r$.
- $\widehat{p}_{a,\ell}$ is predicted size composition for modified
  composition series $a$.

BBRKC uses the four group states male-new, male-old, female-new, and
female-old. Immature crab are not active in this model configuration, but the
RTMB arrays preserve the GMACS group structure.

### B. Population Dynamics

#### Leading Parameters

The male base natural mortality and female multiplier are represented by
leading parameters:

$$
M^0_{\mathrm{male}} = \theta_1,
\qquad
M^0_{\mathrm{female}} = \theta_1 \exp(\theta_2).
$$

Mean recruitment is parameterized on the log scale:

$$
\bar{R} = \exp(\theta_5).
$$

<details>
<summary>R implementation: leading parameter unpacking</summary>

```r
initialize_bbrkc_model_parameters <- function(par, data, predictions = TRUE) {
  theta <- par$theta
  m0 <- c(theta[1], theta[1] * exp(theta[2]))
  model <- list(
    M0 = m0,
    logR0 = theta[3],
    logRini = theta[4],
    logRbar = theta[5],
    logSigmaR = theta[10],
    rho = theta[12]
  )
  ...
}
```

</details>

#### Initial Abundance

GMACS stores the initial numbers-at-size as scaled free parameters. The RTMB
port follows the ADMB convention by setting the first log-scale element to zero
and reading the remaining elements from the active `theta` block:

$$
\log \widetilde{N}_{g,\ell} =
\begin{cases}
0, & (g,\ell) = (1,1),\\
\theta_q, & \text{otherwise}.
\end{cases}
$$

Initial abundance is normalized and scaled by initial recruitment:

$$
N_{g,i_0,1,\ell}
=
\exp(\log R_{\mathrm{init}})
\frac{\exp(\log \widetilde{N}_{g,\ell})}
{\sum_{g'}\sum_{\ell'} \exp(\log \widetilde{N}_{g',\ell'})}.
$$

#### Recruitment

Annual recruitment deviations are additive on the log scale. Sex allocation is
controlled by a logit-scale recruitment proportion:

$$
p_i = \operatorname{logit}^{-1}(\eta_i),
$$

$$
R_{\mathrm{male},i}
=
2 \exp(\log \bar{R} + d_i) p_i,
\qquad
R_{\mathrm{female},i}
=
2 \exp(\log \bar{R} + d_i) (1 - p_i).
$$

Recruitment enters the new-shell group according to the recruitment size
distribution $\rho_{h,\ell}$:

$$
N_{\mathrm{new},h,i,1,\ell}
\leftarrow
N_{\mathrm{new},h,i,1,\ell}
+ R_{h,i}\rho_{h,\ell}.
$$

<details>
<summary>R implementation: recruitment and sex split</summary>

```r
calc_bbrkc_recruits <- function(par, data, model) {
  rec_rows <- years >= controls$rdv_syr & years <= controls$rdv_eyr
  rec_dev[rec_rows] <- par$rec_dev_est[seq_len(sum(rec_rows))]
  logit_rec_prop[rec_rows] <- par$logit_rec_prop_est[seq_len(sum(rec_rows))]

  for (iy in seq_len(n_years)) {
    total_recruitment <- exp(model$logRbar) * nsex
    total_recruitment <- total_recruitment * exp(rec_dev[iy])
    male_prop <- 1 / (1 + exp(-logit_rec_prop[iy]))
    recruits[1, iy] <- total_recruitment * male_prop
    recruits[2, iy] <- total_recruitment * (1 - male_prop)
  }
}
```

</details>

#### Natural Mortality

For sex $h$ and maturity state $m$, natural mortality is

$$
M_{h,m,i,\ell}
=
M^0_h \exp(\delta_{h,b(i)}) Q_{h,m,\ell},
$$

where $\delta_{h,b(i)}$ is the block-specific natural-mortality deviation and
$Q_{h,m,\ell}$ is an optional size-specific multiplier. In the current BBRKC
input, the active size-specific multiplier is effectively one:

$$
Q_{h,m,\ell} = 1.
$$

#### Selectivity, Retention, and Fishing Mortality

The port currently supports the logistic GMACS selectivity form used by the
BBRKC fleets:

$$
s_{k,h,i,\ell}
=
\frac{
\left[1 + \exp\left\{-(L_\ell - x_{50,k,h,i})/\gamma_{k,h,i}\right\}\right]^{-1}
}
{\max_{\ell'} \left[1 + \exp\left\{-(L_{\ell'} - x_{50,k,h,i})/\gamma_{k,h,i}\right\}\right]^{-1}},
$$

where $L_\ell$ is the size-bin midpoint, $x_{50}$ is the size at 50 percent
selection, and $\gamma$ controls the slope. Retention uses the same logistic
form with retention-specific parameters:

$$
r_{k,h,i,\ell}
=
\left[1 + \exp\left\{-(L_\ell - x^{R}_{50,k,h,i})/\gamma^R_{k,h,i}\right\}\right]^{-1}.
$$

The fleet-specific availability multiplier combines capture selectivity,
retention, and the high-grading or discard mortality term $\xi_k$:

$$
V_{k,h,i,\ell}
=
s_{k,h,i,\ell}
\,\left[
r_{k,h,i,\ell}
+ (1-r_{k,h,i,\ell})\xi_k
\right].
$$

Fully selected fleet mortality is

$$
f_{k,h,i,j} = \exp(\log f_{k,h,i,j}),
$$

and total fishing mortality at size is summed over active fleets:

$$
F_{h,i,j,\ell}
=
\sum_k f_{k,h,i,j} V_{k,h,i,\ell}.
$$

<details>
<summary>R implementation: selectivity, retention, and fishing mortality</summary>

```r
calc_bbrkc_fishing_mortality <- function(par, data, model) {
  for (fleet in seq_len(nfleet)) {
    for (sex in seq_len(nsex)) {
      ...
      sel <- exp(model$log_slx_capture[fleet, sex, iy, ]) + 1e-10
      ret <- exp(model$log_slx_retaind[fleet, sex, iy, ]) * slx_nret[sex, fleet]
      vul <- sel * (ret + (1 - ret) * xi)
      F[sex, iy, season, ] <- F[sex, iy, season, ] +
        ft[fleet, sex, iy, season] * vul
      F2[sex, iy, season, ] <- F2[sex, iy, season, ] +
        ft[fleet, sex, iy, season] * sel
    }
  }
}
```

</details>

#### Total Mortality and Survival

Continuous-season total mortality is

$$
Z_{h,m,i,j,\ell}
=
\tau^M_j M_{h,m,i,\ell}
+ F_{h,i,j,\ell},
$$

where $\tau^M_j$ is the proportion of annual natural mortality applied in
season $j$. Survival through the season is

$$
\phi_{h,m,i,j,\ell} = \exp(-Z_{h,m,i,j,\ell}).
$$

For discrete-season accounting, GMACS also stores the equivalent annualized
total mortality used for reporting:

$$
Z^{D}_{h,m,i,j,\ell}
=
\frac{-\log\left[
1 -
\left[
1 - \exp\left\{
-(\tau^M_j M_{h,m,i,\ell} + F_{h,i,j,\ell})
\right\}
\right]
\right]}
\tau_j,
$$

where $\tau_j$ is the modeled season length.

#### Molting and Growth

The molt probability is represented as a declining logistic curve:

$$
P^{\mathrm{molt}}_{h,i,\ell}
=
1 -
\left[
1 + \exp\left\{-\frac{L_\ell-\mu^{\mathrm{molt}}_{h,i}}
{\sigma^{\mathrm{molt}}_{h,i}}\right\}
\right]^{-1}.
$$

Growth is represented by a transition matrix $G_{h,i,\ell,\ell'}$:

$$
G_{h,i,\ell,\ell'} =
\Pr(L_{\ell'} \mid L_\ell,\ h,\ i),
\qquad
\sum_{\ell'} G_{h,i,\ell,\ell'} = 1.
$$

The current RTMB port computes the same transition matrix as ADMB to text-output
precision, using the BBRKC molt-increment controls and the same terminal-bin
normalization.

<details>
<summary>R implementation: molt probability and growth transition</summary>

```r
calc_bbrkc_molting_probability <- function(growth, data) {
  for (sex in seq_len(nsex)) {
    location <- growth$molt_mu[sex]
    scale <- growth$molt_mu[sex] * growth$molt_cv[sex]
    out[sex, ] <- 1 - plogis_gmacs(data$mid_points, location, scale)
  }
  out
}

calc_bbrkc_growth_transition <- function(growth, data) {
  for (sex in seq_len(nsex)) {
    for (from in seq_len(nclass)) {
      cdf <- ad_pgamma(data$size_breaks - data$mid_points[from],
        shape = shape[from], scale = scale
      )
      probs <- diff(cdf)
      out[sex, from, ] <- probs / sum(probs)
    }
  }
  out
}
```

</details>

#### Season and Year Propagation

Let post-mortality abundance be

$$
\widetilde{N}_{g,i,j,\ell}
=
N_{g,i,j,\ell}\phi_{h,m,i,j,\ell}.
$$

For new-shell crab, the nonmolting component moves to old shell and the molting
component remains new shell after growth:

$$
N^{\mathrm{old}}_{h,i,j+1,\ell}
\leftarrow
N^{\mathrm{old}}_{h,i,j+1,\ell}
+ \widetilde{N}^{\mathrm{new}}_{h,i,j,\ell}
\left(1-P^{\mathrm{molt}}_{h,i,\ell}\right),
$$

$$
N^{\mathrm{new}}_{h,i,j+1,\ell'}
\leftarrow
N^{\mathrm{new}}_{h,i,j+1,\ell'}
+ \sum_\ell
\widetilde{N}^{\mathrm{new}}_{h,i,j,\ell}
P^{\mathrm{molt}}_{h,i,\ell}
G_{h,i,\ell,\ell'}.
$$

For old-shell crab, the nonmolting component remains old shell and the molting
component returns to new shell after growth:

$$
N^{\mathrm{old}}_{h,i,j+1,\ell}
\leftarrow
N^{\mathrm{old}}_{h,i,j+1,\ell}
+ \widetilde{N}^{\mathrm{old}}_{h,i,j,\ell}
\left(1-P^{\mathrm{molt}}_{h,i,\ell}\right),
$$

$$
N^{\mathrm{new}}_{h,i,j+1,\ell'}
\leftarrow
N^{\mathrm{new}}_{h,i,j+1,\ell'}
+ \sum_\ell
\widetilde{N}^{\mathrm{old}}_{h,i,j,\ell}
P^{\mathrm{molt}}_{h,i,\ell}
G_{h,i,\ell,\ell'}.
$$

At the terminal season, the propagated state becomes the first season of the
next modeled year.

<details>
<summary>R implementation: population update</summary>

```r
calc_bbrkc_population_numbers <- function(par, data, model) {
  for (iy in seq_len(n_years)) {
    for (season in seq_len(nseason)) {
      for (group in seq_len(n_groups)) {
        survived <- N[group, iy, season, ] * model$S[sex, maturity, iy, season, ]
        molted <- survived * model$molt_probability[sex, ]
        not_molted <- survived * (1 - model$molt_probability[sex, ])
        grown <- as.vector(molted %*% model$growth_transition[sex, , ])
        ...
      }
    }
  }
}
```

</details>

### C. Observation and Prediction Equations

#### Catch Predictions

For catch observation row $r$, let $k(r)$ be the fleet, $i(r)$ the year,
$j(r)$ the season, $H_r$ the included sexes, and $T_r$ the catch type. The
catch-type availability is

$$
A^{T_r}_{k,h,i,\ell} =
\begin{cases}
s_{k,h,i,\ell}, & T_r = \mathrm{capture},\\
s_{k,h,i,\ell} r_{k,h,i,\ell}, & T_r = \mathrm{retained},\\
s_{k,h,i,\ell} (1-r_{k,h,i,\ell}), & T_r = \mathrm{discard}.
\end{cases}
$$

Predicted catch uses the Baranov catch equation:

$$
\widehat{C}_r
=
\sum_{h \in H_r}\sum_m\sum_o\sum_\ell
W^{u_r}_{h,m,\ell}
N_{h,m,o,i(r),j(r),\ell}
f_{k(r),h,i(r),j(r)}
A^{T_r}_{k(r),h,i(r),\ell}
\frac{1-\exp[-Z_{h,m,i(r),j(r),\ell}]}
{Z_{h,m,i(r),j(r),\ell}},
$$

where $u_r$ determines whether the observation is in numbers or biomass:

$$
W^{u_r}_{h,m,\ell} =
\begin{cases}
1, & u_r = \mathrm{numbers},\\
w_{h,m,\ell}, & u_r = \mathrm{biomass}.
\end{cases}
$$

The catch residual currently reported by RTMB is

$$
e^{C}_r = \log(C^{\mathrm{obs}}_r) - \log(\widehat{C}_r).
$$

When GMACS treats catch as effort with analytic catchability, the same predicted
catch vector is used to estimate

$$
\log \widehat{q}_r =
\log(C^{\mathrm{obs}}_r) - \log(\widehat{C}_r).
$$

<details>
<summary>R implementation: catch prediction</summary>

```r
calc_bbrkc_predicted_catch <- function(data, model) {
  for (series in seq_along(data$catch$frames)) {
    for (row in seq_len(nrow(frame))) {
      ...
      nal <- calc_bbrkc_catch_weighted_numbers(
        model, data, sex, maturity, group, year_index, season, units
      )
      mult <- calc_bbrkc_catch_mortality_multiplier(
        model, data, sex, maturity, year_index, season
      )
      pre <- pre + sum(nal * model$ft[fleet, sex, year_index, season] * sel * mult)
      residual <- log(cobs) - log(pre)
    }
  }
}
```

</details>

#### Size-Composition Predictions

For catch-like size-composition rows, the unnormalized predicted count at size
uses the same catch mortality multiplier as the catch equation:

$$
D_{r,\ell}
=
\sum_{h \in H_r}\sum_m\sum_o
N_{h,m,o,i(r),j(r),\ell}
f_{k(r),h,i(r),j(r)}
A^{T_r}_{k(r),h,i(r),\ell}
\frac{1-\exp[-Z_{h,m,i(r),j(r),\ell}]}
{Z_{h,m,i(r),j(r),\ell}}.
$$

For survey-like size-composition rows, the prediction is based on selected
numbers at size:

$$
D_{r,\ell}
=
\sum_{h \in H_r}\sum_m\sum_o
N_{h,m,o,i(r),j(r),\ell}
A^{T_r}_{k(r),h,i(r),\ell}.
$$

After optional tail compression, rows assigned to the same modified composition
series $a$ are accumulated and normalized:

$$
\widehat{p}_{a,\ell}
=
\frac{\sum_{r \in a} D_{r,\ell}}
{\sum_{\ell'}\sum_{r \in a} D_{r,\ell'}}.
$$

The current RTMB report stores this as a flat vector
`model$pre_size_comps`, with dimensions carried by `model$size_comp_rows` and
`model$size_comp_cols`.

<details>
<summary>R implementation: size-composition prediction and aggregation</summary>

```r
calc_bbrkc_predicted_size_compositions <- function(data, model) {
  for (series in seq_len(n_input)) {
    ...
    if (controls$lf_catch_in[series] == 1L) {
      mult <- calc_bbrkc_catch_mortality_multiplier(...)
    } else {
      mult <- rep(1, data$dimensions[["nclass"]])
    }
    dntmp <- dntmp + model$N[group, year_index, season, ] * sel * mult
  }

  for (modified in seq_len(n_modified)) {
    input_series <- which(controls$comp_aggregator == modified)
    row_pred <- do.call(c, row_values)
    pred_rows[[row]] <- row_pred / sum(row_pred)
  }
}
```

</details>

#### Survey Index Predictions

Survey abundance and biomass predictions are active in RTMB. For survey row
$r$, the vulnerable selected abundance or biomass before catchability scaling is

$$
V_r
=
\sum_{h \in H_r}\sum_m\sum_o\sum_\ell
W^{u_r}_{h,m,\ell}
N_{h,m,o,i(r),j(r),\ell}
B_{k(r),h,i(r),\ell},
$$

where $B_{k,h,i,\ell}$ is survey selectivity, optionally including retention or
other survey-specific availability modifiers depending on the GMACS survey
type. If the survey occurs after part of the season has elapsed, the selected
numbers are discounted by $\exp(-t_r Z_{h,m,i,j,\ell})$ before summation.

Predicted index is

$$
\widehat{I}_r = q_{s(r)} V_r,
$$

where $s(r)$ is the survey series. For analytic catchability,

$$
\log \widehat{q}_{k}
=
\frac{
\sum_{r \in k} \omega_r
\left[\log(I^{\mathrm{obs}}_r) - \log(\widehat{I}_r/q_k)\right]
}{
\sum_{r \in k} \omega_r
}.
$$

The BBRKC run currently estimates or fixes survey q directly from the `.pin`
state rather than using analytic q. The reported RTMB `model$pre_cpue` vector is
checked against `#Index_fit_summary` in `Gmacsall.out`.

<details>
<summary>R implementation: survey prediction</summary>

```r
calc_bbrkc_predicted_survey <- function(par, data, model) {
  for (row in seq_len(nrow(survey))) {
    ...
    survey_availability <- if (data$survey$index_type[series] == 2L) {
      sel * ret
    } else {
      sel
    }
    nal <- calc_bbrkc_survey_weighted_numbers(
      model, data, sex, maturity, group, year_index, season, units
    )
    row_v <- row_v + sum(nal * survey_availability * exp(-z_at_time))
  }

  pred <- survey_q_vec[series] * vulnerable_vec[row]
  pre_cpue[[row]] <- pred
  res_cpue[[row]] <- log(survey$obs[row]) - log(pred)
}
```

</details>

### D. Model Fitting and Likelihoods

The current RTMB objective includes the likelihood components whose prediction
vectors have been ported and validated:

$$
\mathcal{L}
=
\mathcal{L}_{\mathrm{catch}}
+ \mathcal{L}_{\mathrm{comp}}
+ \mathcal{L}_{\mathrm{survey}}
+ \mathcal{L}_{\mathrm{recruit}}
+ \mathcal{P}_{\mathrm{weighted}}
+ \mathcal{D}_{\mathrm{prior}}
+ 0_{\mathrm{growth}}
$$

Weighted penalties and prior densities are active in the objective. Growth
likelihood is zero because the current BBRKC data file has no active growth
observations.

<details>
<summary>R implementation: objective assembly</summary>

```r
nloglike <- c(
  catch = sum(catch_nloglike * data$likelihood_emphasis$catch),
  index = sum(index_nloglike * data$likelihood_emphasis$cpue_emphasis),
  length = sum(length_nloglike * data$observed_size_compositions$emphasis),
  recruitment = sum(recruitment_nloglike),
  growth = growth_nloglike
)

nll + sum(nloglike) + sum(weighted_nlog_penalty) + sum(prior_density)
```

</details>

#### Catch Likelihood

For lognormal catch observations,

$$
\sigma^C_r = \sqrt{\log(1 + \mathrm{cv}_{r}^{2})},
\qquad
e^C_r = \log(C^{\mathrm{obs}}_r) - \log(\widehat{C}_r),
$$

and the contribution is

$$
\mathcal{L}_{\mathrm{catch}}
=
\sum_r \lambda^C_r
\left[
\frac{1}{2}\left(\frac{e^C_r}{\sigma^C_r}\right)^2
+ \log(\sigma^C_r\sqrt{2\pi})
\right].
$$

Rows configured as effort observations use the same residual form after applying
the analytic or estimated catchability parameter.

<details>
<summary>R implementation: catch likelihood</summary>

```r
admb_dnorm_nll <- function(x, sd, mean = 0) {
  sum(0.5 * log(2 * pi) + log(sd) + 0.5 * ((x - mean) / sd)^2)
}

catch_likelihood <- function() {
  out <- vector("list", length(data$catch$frames))
  start <- 1L
  for (series in seq_along(data$catch$frames)) {
    frame <- data$catch$frames[[series]]
    rows <- nrow(frame)
    idx <- start:(start + rows - 1L)
    catch_sd <- sqrt(log(1 + frame$cv^2))
    out[[series]] <- admb_dnorm_nll(model$res_catch[idx], catch_sd)
    start <- start + rows
  }
  do.call(c, out)
}
```

</details>

#### Survey Likelihood

For lognormal survey index observations with optional additive process error,

$$
\sigma^I_r
=
\sqrt{\log\left(1 + \mathrm{cv}_{r}^{2}\right)
+ \sigma^2_{\mathrm{add},k(r)}},
$$

$$
e^I_r = \log(I^{\mathrm{obs}}_r) - \log(\widehat{I}_r),
$$

$$
\mathcal{L}_{\mathrm{survey}}
=
\sum_r \lambda^I_r
\left[
\frac{1}{2}\left(\frac{e^I_r}{\sigma^I_r}\right)^2
+ \log(\sigma^I_r\sqrt{2\pi})
\right].
$$

<details>
<summary>R implementation: survey likelihood</summary>

```r
index_likelihood <- function() {
  for (series in seq_along(out)) {
    rows <- which(survey$index == series)
    for (row in rows) {
      cvadd2 <- if (controls$add_cv_links[series] > 0L) {
        log(1 + exp(par$log_add_cv[controls$add_cv_links[series]])^2)
      } else {
        0
      }
      cvobs2 <- log(1 + survey$cv[row]^2) / controls$cpue_lambda[series]
      stdtmp <- sqrt(cvobs2 + cvadd2)
      nll_series <- nll_series + log(stdtmp) +
        0.5 * (model$res_cpue[row] / stdtmp)^2
    }
  }
}
```

</details>

#### Size-Composition Likelihood

For multinomial size compositions with effective sample size $n_a$ and observed
proportions $p^{\mathrm{obs}}_{a,\ell}$,

$$
\mathcal{L}_{\mathrm{comp},a}
=
-\lambda_a n_a
\sum_\ell
p^{\mathrm{obs}}_{a,\ell}
\log(\widehat{p}_{a,\ell}).
$$

GMACS also supports robust and Dirichlet-like composition likelihoods. A
Dirichlet form can be written as

$$
\alpha_{a,\ell} = n_a \widehat{p}_{a,\ell},
$$

$$
\mathcal{L}_{\mathrm{Dirichlet},a}
=
-
\left[
\log\Gamma\left(\sum_\ell \alpha_{a,\ell}\right)
- \sum_\ell \log\Gamma(\alpha_{a,\ell})
+ \sum_\ell
(\alpha_{a,\ell}-1)\log(p^{\mathrm{obs}}_{a,\ell})
\right].
$$

The active BBRKC likelihood type should be ported from the parsed
`size_comp_controls$likelihood_type_in` values rather than hard-coded.

<details>
<summary>R implementation: robust multinomial likelihood used by BBRKC</summary>

```r
robust_multinomial_nll_flat <- function(obs, pred, log_effn) {
  eps <- 1e-08
  nll <- 0
  cols <- ncol(obs)
  a <- 0.1 / cols
  effn <- exp(log_effn)
  for (i in seq_len(nrow(obs))) {
    row_start <- (i - 1L) * cols
    psum <- 0
    for (j in seq_len(cols)) {
      psum <- psum + pred[row_start + j] + eps
    }
    o <- obs[i, ] + eps
    o <- o / sum(o)
    for (j in seq_len(cols)) {
      p <- (pred[row_start + j] + eps) / psum
      v <- a + o[j] * (1 - o[j])
      ell <- 0.5 * ((p - o[j])^2 / v)
      nll <- nll - log(exp(-effn[i] * ell) + 0.01)
      nll <- nll + 0.5 * log(v / effn[i])
    }
  }
  nll
}
```

</details>

#### Recruitment Likelihood

Recruitment deviations are fit through the ADMB `res_recruit` vector. For the
BBRKC no-stock-recruitment option, the first residual is

$$
\epsilon_{i_0}
=
\log R_{\mathrm{male},i_0}
- \log R_0
+ \frac{1}{2}\sigma_R^2.
$$

For later years,

$$
\epsilon_i
=
\log R_{\mathrm{male},i}
- (1-\rho_R)\log \bar{R}
- \rho_R \log R_{\mathrm{male},i-1}
+ \frac{1}{2}\sigma_R^2.
$$

The likelihood contribution is

$$
\mathcal{L}_{\mathrm{recruit}}
=
\sum_i
\left[
\frac{1}{2}\left(\frac{\epsilon_i}{\sigma_R}\right)^2
+ \log(\sigma_R\sqrt{2\pi})
\right].
$$

<details>
<summary>R implementation: recruitment residual and likelihood</summary>

```r
calc_bbrkc_recruitment_residual <- function(par, data, model) {
  sigR <- exp(par$theta[10])
  sig2R <- 0.5 * sigR * sigR
  male_recruits <- model$recruits[1, ]
  out[[1]] <- log(male_recruits[1]) - model$logR0 + sig2R
  for (iy in 2:n_years) {
    out[[iy]] <- log(male_recruits[iy]) -
      (1 - model$rho) * model$logRbar -
      model$rho * log(male_recruits[iy - 1L]) +
      sig2R
  }
  do.call(c, out)
}

recruitment_likelihood <- function() {
  sigR <- exp(par$theta[10])
  out[[1]] <- admb_dnorm_nll(model$res_recruit, sigR)
  out[[3]] <- admb_dnorm_nll(par$logit_rec_prop_est, 2)
  do.call(c, out)
}
```

</details>

#### Growth Likelihood

Growth-increment observations, when active, are fit by the GMACS growth model
using the predicted increment mean and variance:

$$
\Delta L_n \sim
\operatorname{Normal}
\left(
\mu^{\Delta L}_{h(n)}(L_n),
\sigma^{\Delta L}_{h(n)}(L_n)
\right).
$$

The corresponding negative log-likelihood is

$$
\mathcal{L}_{\mathrm{growth}}
=
\sum_n
\left[
\frac{1}{2}
\left(
\frac{\Delta L^{\mathrm{obs}}_n
- \mu^{\Delta L}_{h(n)}(L_n)}
{\sigma^{\Delta L}_{h(n)}(L_n)}
\right)^2
+ \log\left(\sigma^{\Delta L}_{h(n)}(L_n)\sqrt{2\pi}\right)
\right].
$$

Growth likelihood activation should occur after the deterministic growth
transition matrix remains matched to ADMB under the same input data.

### E. Penalties and Priors

The GMACS objective includes penalties and priors in addition to observation
likelihoods. The RTMB port now computes the raw `nlogPenalty` vector and adds
the weighted penalty contribution to the objective:

$$
\mathcal{P}_{\mathrm{weighted}}
=
\sum_{b=1}^{13} w_b \mathcal{P}_b.
$$

The nonzero BBRKC raw penalty components are:

$$
\mathcal{P}_1
=
\sum_k
\omega^{F}_{k,1}\overline{d}^{\,2}_{F,k}
+ \omega^{F}_{k,2}\overline{d}^{\,2}_{F\mathrm{off},k},
$$

$$
\mathcal{P}_3
=
\sum_b
\left[
\frac{1}{2}\left(\frac{\delta^M_b}{\sigma_M}\right)^2
+ \log(\sigma_M\sqrt{2\pi})
\right],
$$

$$
\mathcal{P}_6
=
\sum_i
\left[
\frac{1}{2}\left(d_i-d_{i-1}\right)^2
+ \log(\sqrt{2\pi})
\right],
$$

$$
\mathcal{P}_7
=
\left[
\log\left(\sum_i R_{\mathrm{female},i}\right)
- \log\left(\sum_i R_{\mathrm{male},i}\right)
\right]^2,
$$

and the initial abundance smoothness penalty is

$$
\mathcal{P}_{10}
=
\sum_g\sum_{\ell=2}^{L}
\left[
\frac{1}{2}
\left(\log \widetilde{N}_{g,\ell}
- \log \widetilde{N}_{g,\ell-1}\right)^2
+ \log(\sqrt{2\pi})
\right].
$$

The raw RTMB penalty vector matches the ADMB `nlogPenalty` block to printed
precision. The weighted contribution is smaller than the raw sum because the
BBRKC control file sets several penalty-emphasis weights to zero.

<details>
<summary>R implementation: penalty blocks</summary>

```r
calc_bbrkc_penalties <- function(par, data, model, zero_ad) {
  out <- vector("list", 13)
  for (i in seq_along(out)) out[[i]] <- zero_ad

  log_fdev <- split_parameter_by_fleet(par$log_fdev, data$fdev_indices)
  log_fdov <- split_parameter_by_fleet(par$log_fdov, data$fdov_indices)
  for (fleet in seq_along(log_fdev)) {
    out[[1]] <- out[[1]] +
      data$likelihood_emphasis$fdev_penalty[fleet, 1] * mean(log_fdev[[fleet]])^2
    out[[1]] <- out[[1]] +
      data$likelihood_emphasis$fdev_penalty[fleet, 2] * mean(log_fdov[[fleet]])^2
  }

  out[[3]] <- admb_dnorm_nll(par$m_dev_est, data$m_controls$mdev_sd)
  out[[6]] <- first_difference_nll(model$rec_dev, 1)
  out[[7]] <- (log(sum(model$recruits[2, ])) - log(sum(model$recruits[1, ])))^2
  for (group in seq_len(nrow(model$logN0))) {
    out[[10]] <- out[[10]] + first_difference_nll(model$logN0[group, ], 1)
  }

  do.call(c, out)
}

weighted_nlog_penalty <- nlog_penalty * data$likelihood_emphasis$penalty
```

</details>

Parameter priors enter as negative log densities:

$$
\mathcal{D}_{\mathrm{prior}}
=
-\sum_b \log \pi_b(\psi_b),
$$

where $\psi_b$ is an active parameter block and $\pi_b$ is the corresponding
prior density specified by GMACS. The RTMB port preserves the ADMB
`priorDensity` vector layout, including zero entries for active parameters that
intentionally have no prior.

The implemented prior forms are:

$$
\mathcal{D}_{\mathrm{uniform}}(x; a,b) = \log(b-a),
$$

$$
\mathcal{D}_{\mathrm{normal}}(x; \mu,\sigma)
=
\frac{1}{2}\log(2\pi) + \log(\sigma)
+ \frac{1}{2}\left(\frac{x-\mu}{\sigma}\right)^2,
$$

$$
\mathcal{D}_{\mathrm{lognormal}}(x; m,\sigma)
=
\frac{1}{2}\log(2\pi) + \log(\sigma) + \log(x)
+ \frac{1}{2}\left(\frac{\log(x)-\log(m)}{\sigma}\right)^2,
$$

with beta and gamma negative log densities available for configurations that
activate those prior types. Selectivity priors are applied to
$\exp(\log s)$, matching GMACS. Additional-CV priors are applied to
$\exp(\log \sigma_{\mathrm{add}})$, also matching GMACS.

<details>
<summary>R implementation: prior-density blocks</summary>

```r
admb_prior_nll <- function(type, x, p1, p2) {
  switch(
    as.character(as.integer(type)),
    "0" = log(p2 - p1),
    "1" = admb_dnorm_nll(x, p2, p1),
    "2" = 0.5 * log(2 * pi) + log(p2) + log(x) +
      0.5 * ((log(x) - log(p1)) / p2)^2,
    "3" = -(lgamma(p1 + p2) - lgamma(p1) - lgamma(p2) +
      (p1 - 1) * log(x) + (p2 - 1) * log(1 - x)),
    "4" = -(p1 * log(p2) - lgamma(p1) +
      (p1 - 1) * log(x) - p2 * x)
  )
}

calc_bbrkc_prior_density <- function(par, data, zero_ad) {
  out <- rep(0, length(data$admb_reference$prior_density))
  iprior <- 1L

  ## Fill active theta, growth, selectivity, M, q, add-CV, and mature-M priors.
  ## Advance iprior over active parameters that have no prior, preserving ADMB
  ## priorDensity vector layout.
  ...

  out + zero_ad
}
```

</details>

The parser reads the ADMB `nloglike`, `nlogPenalty`, and `priorDensity` summary
blocks from `Gmacsall.out`; all currently active BBRKC objective components are
checked against those component-level regression targets.

`audit_bbrkc_admb_components()` builds the current regression table from an
RTMB report object and the parsed ADMB reference blocks. It aggregates the
likelihood rows to match the BBRKC objective components and compares the raw
penalty and prior vectors element by element.

## Appendix: Results Figures

The figures below are generated directly from the current RTMB population
state initialized at the BBRKC fitted `.pin` values. They are not a refit; they
are quick visual diagnostics that the port produces interpretable annual
trajectories for biomass, recruitment, and natural mortality.

```{r}
#| label: results-setup
#| include: false
has_bbrkc_inputs <- all(file.exists(file.path(
  "examples", "BBRKC",
  c("gmacs.dat", "gmacs.pin", "Gmacsall.out", "gmacs.rep")
)))

if (has_bbrkc_inputs) {
source("R/read_gmacs_bbrkc.R")
source("R/gmacs_rtmb_nll.R")

inputs <- read_bbrkc_inputs("examples/BBRKC")
rtmb_data <- make_gmacs_rtmb_data(inputs)
rtmb_par <- make_gmacs_parameter_list(inputs)
rtmb_model <- initialize_bbrkc_model_parameters(rtmb_par, rtmb_data)

years <- rtmb_data$years
n_years <- length(years)
groups <- rtmb_model$groups
}
```

```{r}
#| label: missing-results-inputs
#| echo: false
#| results: asis
if (!has_bbrkc_inputs) {
  cat(
    "The executable results figures require `examples/BBRKC/gmacs.dat`, ",
    "`gmacs.pin`, `Gmacsall.out`, and `gmacs.rep`. Those local model inputs ",
    "are not committed to the repository, so the GitHub Pages render shows ",
    "the figure code without executing it."
  )
}
```

### Biomass

Annual biomass is summarized from the first season of each model year as
numbers-at-size multiplied by mean weight-at-size, summed over sex, maturity,
shell condition, and size class:

$$
B_y =
\sum_g\sum_\ell
N_{g,y,1,\ell}\,
w_{\mathrm{sex}(g),\mathrm{mat}(g),y,\ell}.
$$

```{r}
#| label: fig-rtmb-biomass
#| fig-cap: "BBRKC biomass trajectory from the current RTMB population state."
#| code-fold: true
#| eval: !expr has_bbrkc_inputs
biomass <- numeric(n_years)
for (iy in seq_len(n_years)) {
  for (ig in seq_len(nrow(groups))) {
    sex <- groups$sex[ig]
    maturity <- groups$maturity[ig]
    biomass[iy] <- biomass[iy] +
      sum(rtmb_model$N[ig, iy, 1, ] * rtmb_data$mean_wt[sex, maturity, iy, ])
  }
}

plot(
  years, biomass, type = "l", lwd = 2, col = "#1b6da8",
  xlab = "Year", ylab = "Biomass (model units)", bty = "l"
)
grid(col = "grey85")
```

### Recruitment

Recruitment is summarized by sex and in total from the ported GMACS
recruitment block:

$$
R_y = \sum_s R_{s,y}.
$$

```{r}
#| label: fig-rtmb-recruitment
#| fig-cap: "BBRKC recruitment trajectory from the current RTMB recruitment block."
#| code-fold: true
#| eval: !expr has_bbrkc_inputs
recruits <- rtmb_model$recruits[, seq_len(n_years), drop = FALSE]
recruitment_total <- colSums(recruits)

matplot(
  years, t(recruits), type = "l", lty = 1, lwd = 1.5,
  col = c("#386cb0", "#fdb462"),
  xlab = "Year", ylab = "Recruitment (numbers)", bty = "l"
)
lines(years, recruitment_total, lwd = 2.5, col = "#222222")
grid(col = "grey85")
legend(
  "topright", bty = "n", lty = 1, lwd = c(1.5, 1.5, 2.5),
  col = c("#386cb0", "#fdb462", "#222222"),
  legend = c("Female", "Male", "Total")
)
```

### Natural Mortality

Natural mortality is plotted as an annual size-class mean by sex for the
available mortality maturity stratum:

$$
\overline{M}_{s,y}
=
\frac{1}{L}
\sum_\ell M_{s,m^*,y,\ell}.
$$

```{r}
#| label: fig-rtmb-natural-mortality
#| fig-cap: "Mean natural mortality by sex from the current RTMB mortality block."
#| code-fold: true
#| eval: !expr has_bbrkc_inputs
n_sex <- rtmb_data$dimensions[["nsex"]]
maturity_index <- min(2L, dim(rtmb_model$M)[2])
mean_m <- matrix(NA_real_, nrow = n_years, ncol = n_sex)
for (sex in seq_len(n_sex)) {
  mean_m[, sex] <- rowMeans(rtmb_model$M[sex, maturity_index, seq_len(n_years), ])
}

matplot(
  years, mean_m, type = "l", lty = 1, lwd = 2,
  col = c("#386cb0", "#fdb462"),
  xlab = "Year", ylab = "Mean M", bty = "l"
)
grid(col = "grey85")
legend(
  "topright", bty = "n", lty = 1, lwd = 2,
  col = c("#386cb0", "#fdb462"),
  legend = c("Female", "Male")
)
```
