---
title: "Snow Crab GMACS Input and ADMB Fit Notes"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
execute:
  echo: true
  warning: false
  message: false
---

## Purpose

This document is the Snow crab parallel to the BBRKC and Tanner pages. The Snow
example in `examples/snow` includes GMACS input files plus ADMB output reports,
so this page can summarize both parsed inputs and observed-vs-predicted ADMB
fit diagnostics.

The current RTMB objective implementation is still BBRKC-specific. The Snow
work here establishes parser coverage and report targets for a future Snow RTMB
objective.

## Files

| File | Role |
|---|---|
| `examples/snow/gmacs.dat` | GMACS main run file. |
| `examples/snow/snow_21.dat` | Snow crab data file. |
| `examples/snow/snow_21.ctl` | Snow crab control file. |
| `examples/snow/snow_21.prj` | Snow crab projection file. |
| `examples/snow/gmacs.pin` | Fitted parameter state. |
| `examples/snow/Gmacsall.out` | ADMB likelihood and fit-summary output. |
| `examples/snow/gmacs.rep` | ADMB report output. |

## Reader

```{r}
source("R/read_gmacs_bbrkc.R")
source("R/gmacs_rtmb_nll.R")
source("R/read_gmacs_snow.R")

snow <- read_snow_inputs("examples/snow")
validate_snow_inputs(snow)
```

## Input Summary

```{r}
dims <- snow$data$dimensions
data.frame(
  quantity = names(dims),
  value = as.integer(dims),
  row.names = NULL
)
```

```{r}
data.frame(
  component = c("catch", "survey", "size composition"),
  rows = c(
    sum(snow$data$catch$rows),
    nrow(snow$data$survey$data),
    sum(snow$data$size_comp$rows)
  )
)
```

## ADMB Report Summary

```{r}
snow_likelihood_summary(snow)
```

```{r}
data.frame(
  block = c("catch fit", "index fit", "size fit", "overall summary"),
  rows = c(
    nrow(snow$admb_catch_fit),
    nrow(snow$admb_index_fit),
    nrow(snow$admb_size_fit),
    nrow(snow$admb_overall_summary)
  )
)
```

## Index Fit

Observed survey indices are plotted against ADMB predicted values by series.

```{r}
#| label: fig-snow-index-fit
#| fig-cap: "Observed and predicted Snow crab survey indices by series from ADMB output."
#| code-fold: true
index_fit <- snow$admb_index_fit
series_ids <- sort(unique(index_fit$series))
panel_rows <- ceiling(length(series_ids) / 2)
panel_slots <- panel_rows * 2
old_par <- par(mar = c(3, 3, 2, 1), oma = c(0, 0, 0, 0))
layout(
  rbind(matrix(seq_len(panel_slots), ncol = 2, byrow = TRUE), rep(panel_slots + 1, 2)),
  heights = c(rep(1, panel_rows), 0.16)
)
for (series in series_ids) {
  x <- index_fit[index_fit$series == series, ]
  ylim <- range(c(x$obs, x$predicted), finite = TRUE)
  plot(x$year, x$obs, pch = 16, col = "#1b6da8", ylim = ylim,
       xlab = "", ylab = "", main = paste("Series", series), bty = "l")
  line_keys <- unique(x[c("fleet", "season", "sex", "maturity")])
  for (line_i in seq_len(nrow(line_keys))) {
    keep <- rep(TRUE, nrow(x))
    for (key_name in names(line_keys)) {
      keep <- keep & x[[key_name]] == line_keys[[key_name]][line_i]
    }
    line_x <- x[keep, ]
    line_x <- line_x[order(line_x$year), ]
    lines(line_x$year, line_x$predicted, lwd = 2, col = "#c43c39")
    points(line_x$year, line_x$predicted, pch = 1, col = "#c43c39")
  }
  grid(col = "grey88")
}
if (length(series_ids) < panel_slots) {
  for (unused in seq_len(panel_slots - length(series_ids))) {
    plot.new()
  }
}
par(mar = c(0, 0, 0, 0))
plot.new()
legend("center", horiz = TRUE, bty = "n",
       pch = c(16, NA), lty = c(NA, 1), lwd = c(NA, 2),
       col = c("#1b6da8", "#c43c39"), legend = c("Observed", "Predicted"))
layout(1)
par(old_par)
```

## Size-Composition Fit

The panels below compare observed and predicted size compositions for recent
rows from the first four modified composition series.

```{r}
#| label: fig-snow-size-composition-fit
#| fig-cap: "Observed and predicted Snow crab size compositions for selected recent rows."
#| code-fold: true
size_fit <- snow$admb_size_fit
size_keys <- c("modified_series", "year", "fleet", "season", "sex", "type", "shell", "maturity")
size_key_rows <- unique(size_fit[size_fit$vector == "obs", size_keys])
size_key_rows <- size_key_rows[order(size_key_rows$modified_series, size_key_rows$year), ]
series_ids <- sort(unique(size_key_rows$modified_series))[1:4]
old_par <- par(mar = c(3, 3, 2, 1), oma = c(0, 0, 0, 0))
layout(rbind(matrix(1:4, ncol = 2, byrow = TRUE), c(5, 5)), heights = c(1, 1, 0.16))
for (series in series_ids) {
  candidates <- size_key_rows[size_key_rows$modified_series == series, ]
  key <- candidates[nrow(candidates), ]
  key_match <- rep(TRUE, nrow(size_fit))
  for (key_name in size_keys) {
    key_match <- key_match & size_fit[[key_name]] == key[[key_name]]
  }
  obs <- size_fit[key_match & size_fit$vector == "obs", c("size_bin", "value")]
  pred <- size_fit[key_match & size_fit$vector == "pred", c("size_bin", "value")]
  x <- merge(obs, pred, by = "size_bin", suffixes = c("_observed", "_predicted"))
  x <- x[order(x$size_bin), ]
  observed <- x$value_observed / sum(x$value_observed)
  predicted <- x$value_predicted / sum(x$value_predicted)
  ylim <- range(c(observed, predicted), finite = TRUE)
  mids <- barplot(observed, col = "#1b6da8", border = NA,
       ylim = ylim, xlab = "", ylab = "",
       main = paste("Series", series, key$year), bty = "l")
  axis(1, at = mids, labels = x$size_bin)
  lines(mids, predicted, lwd = 2, col = "#c43c39")
  grid(col = "grey88")
}
if (length(series_ids) < 4) {
  for (unused in seq_len(4 - length(series_ids))) {
    plot.new()
  }
}
par(mar = c(0, 0, 0, 0))
plot.new()
legend("center", horiz = TRUE, bty = "n",
       fill = c("#1b6da8", NA), border = c(NA, NA),
       lty = c(NA, 1), lwd = c(NA, 2), col = c(NA, "#c43c39"),
       legend = c("Observed", "Predicted"))
layout(1)
par(old_par)
```
