---
title: "Snow Crab GMACS Input and ADMB Fit Notes"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
execute:
  echo: true
  warning: false
  message: false
---

## Purpose

This document is the Snow crab parallel to the BBRKC and Tanner pages. The Snow
example in `examples/snow` includes GMACS input files plus ADMB output reports,
so this page can summarize both parsed inputs and observed-vs-predicted ADMB
fit diagnostics.

The current RTMB objective implementation is still BBRKC-specific. The Snow
work here establishes parser coverage and report targets for a future Snow RTMB
objective.

## Files

| File | Role |
|---|---|
| `examples/snow/gmacs.dat` | GMACS main run file. |
| `examples/snow/snow_21.dat` | Snow crab data file. |
| `examples/snow/snow_21.ctl` | Snow crab control file. |
| `examples/snow/snow_21.prj` | Snow crab projection file. |
| `examples/snow/gmacs.pin` | Fitted parameter state. |
| `examples/snow/Gmacsall.out` | ADMB likelihood and fit-summary output. |
| `examples/snow/gmacs.rep` | ADMB report output. |

## Reader

```{r}
source("R/read_gmacs_bbrkc.R")
source("R/gmacs_rtmb_nll.R")
source("R/read_gmacs_snow.R")

snow <- read_snow_inputs("examples/snow")
validate_snow_inputs(snow)
```

## Input Summary

```{r}
dims <- snow$data$dimensions
data.frame(
  quantity = names(dims),
  value = as.integer(dims),
  row.names = NULL
)
```

```{r}
data.frame(
  component = c("catch", "survey", "size composition"),
  rows = c(
    sum(snow$data$catch$rows),
    nrow(snow$data$survey$data),
    sum(snow$data$size_comp$rows)
  )
)
```

## ADMB Report Summary

```{r}
snow_likelihood_summary(snow)
```

```{r}
data.frame(
  block = c("catch fit", "index fit", "size fit", "overall summary"),
  rows = c(
    nrow(snow$admb_catch_fit),
    nrow(snow$admb_index_fit),
    nrow(snow$admb_size_fit),
    nrow(snow$admb_overall_summary)
  )
)
```

## Index Fit

Observed survey indices are plotted against ADMB predicted values by series.

```{r}
#| label: fig-snow-index-fit
#| fig-cap: "Observed and predicted Snow crab survey indices by series from ADMB output."
#| code-fold: true
index_fit <- snow$admb_index_fit
series_ids <- sort(unique(index_fit$series))
old_par <- par(mfrow = c(ceiling(length(series_ids) / 2), 2), mar = c(3, 3, 2, 1), oma = c(1, 1, 0, 0))
for (series in series_ids) {
  x <- index_fit[index_fit$series == series, ]
  ylim <- range(c(x$obs, x$predicted), finite = TRUE)
  plot(x$year, x$obs, pch = 16, col = "#1b6da8", ylim = ylim,
       xlab = "", ylab = "", main = paste("Series", series), bty = "l")
  lines(x$year, x$predicted, lwd = 2, col = "#c43c39")
  grid(col = "grey88")
}
par(old_par)
legend("bottom", inset = -0.02, xpd = NA, horiz = TRUE, bty = "n",
       pch = c(16, NA), lty = c(NA, 1), lwd = c(NA, 2),
       col = c("#1b6da8", "#c43c39"), legend = c("Observed", "Predicted"))
```

## Size-Composition Fit

The panels below compare observed and predicted size compositions for recent
rows from the first four modified composition series.

```{r}
#| label: fig-snow-size-composition-fit
#| fig-cap: "Observed and predicted Snow crab size compositions for selected recent rows."
#| code-fold: true
size_fit <- snow$admb_size_fit
series_ids <- sort(unique(size_fit$modified_series))[1:4]
old_par <- par(mfrow = c(2, 2), mar = c(3, 3, 2, 1), oma = c(1, 1, 0, 0))
for (series in series_ids) {
  x <- size_fit[size_fit$modified_series == series, ]
  years <- sort(unique(x$year))
  year <- tail(years, 1)
  obs <- x[x$year == year & x$vector == "obs", ]
  pred <- x[x$year == year & x$vector == "pred", ]
  obs <- obs[order(obs$size_bin), ]
  pred <- pred[order(pred$size_bin), ]
  observed <- obs$value / sum(obs$value)
  predicted <- pred$value / sum(pred$value)
  ylim <- range(c(observed, predicted), finite = TRUE)
  plot(obs$size_bin, observed, type = "h", lwd = 3, col = "#1b6da8",
       ylim = ylim, xlab = "", ylab = "",
       main = paste("Series", series, year), bty = "l")
  lines(pred$size_bin, predicted, lwd = 2, col = "#c43c39")
  grid(col = "grey88")
}
par(old_par)
legend("bottom", inset = -0.02, xpd = NA, horiz = TRUE, bty = "n",
       lty = c(1, 1), lwd = c(3, 2), col = c("#1b6da8", "#c43c39"),
       legend = c("Observed", "Predicted"))
```
