---
title: "Tanner Crab GMACS Input Notes"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    lightbox: true
execute:
  echo: true
  warning: false
  message: false
---

## Purpose

This document is the Tanner crab parallel to `bbrkc_rtmb_port_steps.qmd`.
The current Tanner work is at the input-orientation stage: the repository now
has a Tanner-specific reader for the available GMACS files in `examples/tanners`.
This page also checks for Tanner ADMB report outputs and reports which parity
targets are available for future RTMB objective and deterministic checks.

## Files

The Tanner reader uses:

| File | Role |
|---|---|
| `examples/tanners/gmacs_26_22_03d5_aEffExp1.Run1.dat` | GMACS main run file. |
| `examples/tanners/TannerCrab_Data202509.EffExp1.dat` | Tanner data file. |
| `examples/tanners/TannerCrab_26_22_03d5_a.Run1.ctl` | Tanner control file. |
| `examples/tanners/TannerCrab_26_22_03d5_a.prj` | Tanner projection file. |
| `examples/tanners/gmacs_26_22_03d5_aEffExp1.par` | Fitted parameter table. |
| `examples/tanners/Gmacsall.out` or Tanner equivalent | Optional ADMB likelihood and fit-summary output. |
| `examples/tanners/gmacs.rep` or Tanner equivalent | Optional ADMB report output for deterministic blocks. |

Optional ADMB parity targets are discovered if they are added to the same
folder. Canonical names are `Gmacsall.out` and `gmacs.rep`; the reader also
checks common Tanner-specific variants such as
`gmacs_26_22_03d5_aEffExp1.Gmacsall.out` and
`gmacs_26_22_03d5_aEffExp1.rep`.

## Current Reader

The Tanner files use a different GMACS text layout from the BBRKC example.
`R/read_gmacs_tanner.R` therefore reads the Tanner file set directly instead of
forcing these inputs through BBRKC-specific parser assumptions.

```{r}
source("R/read_gmacs_bbrkc.R")
source("R/read_gmacs_tanner.R")

tanner <- read_tanner_inputs("examples/tanners")
validate_tanner_inputs(tanner)
has_tanner_index_fit <- !is.null(tanner$admb_reference$index_fit)
has_tanner_size_fit <- !is.null(tanner$admb_reference$size_fit)
```

```{r}
data.frame(
  report = names(tanner$admb_reference$files),
  path = unlist(tanner$admb_reference$files),
  available = file.exists(unlist(tanner$admb_reference$files)),
  row.names = NULL
)
```

Candidate ADMB report names checked by the reader:

```{r}
candidates <- tanner_admb_report_candidates("examples/tanners")
data.frame(
  report_type = rep(names(candidates), lengths(candidates)),
  candidate = basename(unlist(candidates)),
  present = file.exists(unlist(candidates)),
  row.names = NULL
)
```

## Input Summary

```{r}
dims <- tanner$data$dimensions
data.frame(
  quantity = names(dims),
  value = as.integer(dims),
  row.names = NULL
)
```

```{r}
data.frame(
  component = c("catch", "survey", "size composition", "growth", "maturity"),
  dataframes = c(
    tanner$data$catch$n,
    tanner$data$survey$n,
    tanner$data$size_comp$n,
    NA_integer_,
    NA_integer_
  ),
  rows = c(
    sum(tanner$data$catch$rows),
    sum(tanner$data$survey$rows),
    sum(tanner$data$size_comp$rows),
    tanner$data$growth$n,
    tanner$data$maturity$n
  )
)
```

## Parameter Summary

```{r}
data.frame(
  parameter_rows = nrow(tanner$parameters),
  active_estimated_parameters = sum(tanner$parameters$active),
  min_estimated_index = min(tanner$parameters$estimated_index, na.rm = TRUE),
  max_estimated_index = max(tanner$parameters$estimated_index, na.rm = TRUE)
)
```

```{r}
head(tanner$parameters[tanner$parameters$active, c(
  "parameter_index", "estimated_index", "value", "label"
)], 12)
```

## ADMB Parity Report Check

The Tanner reader validates:

- model dimensions and size breaks;
- fishery and survey names;
- catch, survey, and size-composition dataframe counts;
- growth and maturity observation counts;
- fitted parameter table extraction from the `.par` file.

The same style of ADMB target extraction used for BBRKC is wired for Tanner
when report outputs are present:

- `read_admb_likelihood_reference()` for `nloglike`, `nlogPenalty`, and
  `priorDensity`;
- `read_admb_catch_fit_summary()`, `read_admb_index_fit_summary()`, and
  `read_admb_size_fit_summary()` for observed/predicted fit summaries;
- `read_admb_selectivity_block()` for `slx_capture`, `slx_retaind`, and
  `slx_discard` from the report file.

```{r}
admb_ref <- tanner$admb_reference
data.frame(
  target = c(
    "nloglike",
    "nlogPenalty",
    "priorDensity",
    "catch fit summary",
    "index fit summary",
    "size fit summary",
    "slx_capture",
    "slx_retaind",
    "slx_discard"
  ),
  source_file = c(
    rep(admb_ref$files$all, 6),
    rep(admb_ref$files$report, 3)
  ),
  parsed_rows = c(
    if (!is.null(admb_ref$likelihood)) length(admb_ref$likelihood$nloglike) else 0L,
    if (!is.null(admb_ref$likelihood)) length(admb_ref$likelihood$nlog_penalty) else 0L,
    if (!is.null(admb_ref$likelihood)) length(admb_ref$likelihood$prior_density) else 0L,
    if (!is.null(admb_ref$catch_fit)) nrow(admb_ref$catch_fit) else 0L,
    if (!is.null(admb_ref$index_fit)) nrow(admb_ref$index_fit) else 0L,
    if (!is.null(admb_ref$size_fit)) nrow(admb_ref$size_fit) else 0L,
    if (!is.null(admb_ref$selectivity)) nrow(admb_ref$selectivity$capture) else 0L,
    if (!is.null(admb_ref$selectivity)) nrow(admb_ref$selectivity$retained) else 0L,
    if (!is.null(admb_ref$selectivity)) nrow(admb_ref$selectivity$discard) else 0L
  ),
  available = c(
    rep(!is.null(admb_ref$likelihood), 3),
    !is.null(admb_ref$catch_fit),
    !is.null(admb_ref$index_fit),
    !is.null(admb_ref$size_fit),
    rep(!is.null(admb_ref$selectivity), 3)
  ),
  row.names = NULL
)
```

If the parsed row counts are zero, add `Gmacsall.out` and `gmacs.rep`, or one
of the Tanner-equivalent names listed above, to `examples/tanners`. Once those
files are present, this page will expose the ADMB likelihood, fit, and
selectivity targets needed to build Tanner parity checks.

```{r}
if (!has_tanner_index_fit || !has_tanner_size_fit) {
  data.frame(
    diagnostic = c("index observed-vs-predicted", "size observed-vs-predicted"),
    status = c(
      if (has_tanner_index_fit) "available" else "missing ADMB Index_fit_summary",
      if (has_tanner_size_fit) "available" else "missing ADMB Size_fit_summary"
    ),
    required_file = "Tanner Gmacsall.out or equivalent",
    row.names = NULL
  )
}
```

## Index Diagnostics

When Tanner ADMB fit summaries are present, the figure below shows observed and
predicted index fits from `Gmacsall.out`. In the current repository, Tanner
ADMB report output is not present, so this figure is observed-only.

```{r}
#| label: fig-tanner-index-observations
#| fig-cap: "Tanner crab index diagnostics; ADMB predictions are shown when report output is available."
#| code-fold: true
if (!is.null(tanner$admb_reference$index_fit)) {
  index_fit <- tanner$admb_reference$index_fit
  series_ids <- sort(unique(index_fit$series))
  panel_rows <- ceiling(length(series_ids) / 2)
  panel_slots <- panel_rows * 2
  old_par <- par(mar = c(3, 3, 2, 1), oma = c(0, 0, 0, 0))
  layout(
    rbind(matrix(seq_len(panel_slots), ncol = 2, byrow = TRUE), rep(panel_slots + 1, 2)),
    heights = c(rep(1, panel_rows), 0.16)
  )
  for (series in series_ids) {
    x <- index_fit[index_fit$series == series, ]
    ylim <- range(c(x$obs, x$predicted), finite = TRUE)
    plot(x$year, x$obs, pch = 16, col = "#1b6da8", ylim = ylim,
         xlab = "", ylab = "", main = paste("Series", series), bty = "l")
    line_keys <- unique(x[c("fleet", "season", "sex", "maturity")])
    for (line_i in seq_len(nrow(line_keys))) {
      keep <- rep(TRUE, nrow(x))
      for (key_name in names(line_keys)) {
        keep <- keep & x[[key_name]] == line_keys[[key_name]][line_i]
      }
      line_x <- x[keep, ]
      line_x <- line_x[order(line_x$year), ]
      lines(line_x$year, line_x$predicted, lwd = 2, col = "#c43c39")
      points(line_x$year, line_x$predicted, pch = 1, col = "#c43c39")
    }
    grid(col = "grey88")
  }
  if (length(series_ids) < panel_slots) {
    for (unused in seq_len(panel_slots - length(series_ids))) {
      plot.new()
    }
  }
  par(mar = c(0, 0, 0, 0))
  plot.new()
  legend("center", horiz = TRUE, bty = "n",
         pch = c(16, NA), lty = c(NA, 1), lwd = c(NA, 2),
         col = c("#1b6da8", "#c43c39"), legend = c("Observed", "Predicted"))
  layout(1)
  par(old_par)
} else {
  index_obs <- tanner$data$survey_observations
  series_ids <- sort(unique(index_obs$series))
  old_par <- par(mfrow = c(ceiling(length(series_ids) / 2), 2), mar = c(3, 3, 2, 1), oma = c(0, 0, 2, 0))
  for (series in series_ids) {
    x <- index_obs[index_obs$series == series, ]
    plot(x$year, x$obs, type = "b", pch = 16, col = "#1b6da8",
         xlab = "", ylab = "", main = paste(x$fleet[1], x$sex[1], x$maturity[1]),
         bty = "l")
    grid(col = "grey88")
  }
  mtext("Observed only: Tanner ADMB fit summaries are not present", outer = TRUE, line = 0.4)
  par(old_par)
}
```

## Size-Composition Diagnostics

When Tanner ADMB size-fit summaries are present, the panels compare observed
and predicted compositions from `Gmacsall.out`. Otherwise they show observed
input compositions for recent rows from the first four size-composition
dataframes. In the current repository, Tanner ADMB report output is not
present, so this figure is observed-only.

```{r}
#| label: fig-tanner-size-composition-observations
#| fig-cap: "Tanner crab size-composition diagnostics; ADMB predictions are shown when report output is available."
#| code-fold: true
if (!is.null(tanner$admb_reference$size_fit)) {
  size_fit <- tanner$admb_reference$size_fit
  size_keys <- c("modified_series", "year", "fleet", "season", "sex", "type", "shell", "maturity")
  size_key_rows <- unique(size_fit[size_fit$vector == "obs", size_keys])
  size_key_rows <- size_key_rows[order(size_key_rows$modified_series, size_key_rows$year), ]
  series_ids <- sort(unique(size_key_rows$modified_series))[1:4]
  old_par <- par(mar = c(3, 3, 2, 1), oma = c(0, 0, 0, 0))
  layout(rbind(matrix(1:4, ncol = 2, byrow = TRUE), c(5, 5)), heights = c(1, 1, 0.16))
  for (series in series_ids) {
    candidates <- size_key_rows[size_key_rows$modified_series == series, ]
    key <- candidates[nrow(candidates), ]
    key_match <- rep(TRUE, nrow(size_fit))
    for (key_name in size_keys) {
      key_match <- key_match & size_fit[[key_name]] == key[[key_name]]
    }
    obs <- size_fit[key_match & size_fit$vector == "obs", c("size_bin", "value")]
    pred <- size_fit[key_match & size_fit$vector == "pred", c("size_bin", "value")]
    x <- merge(obs, pred, by = "size_bin", suffixes = c("_observed", "_predicted"))
    x <- x[order(x$size_bin), ]
    observed <- x$value_observed / sum(x$value_observed)
    predicted <- x$value_predicted / sum(x$value_predicted)
    ylim <- range(c(observed, predicted), finite = TRUE)
    mids <- barplot(observed, col = "#1b6da8", border = NA,
         ylim = ylim, xlab = "", ylab = "",
         main = paste("Series", series, key$year), bty = "l")
    axis(1, at = mids, labels = x$size_bin)
    lines(mids, predicted, lwd = 2, col = "#c43c39")
    grid(col = "grey88")
  }
  if (length(series_ids) < 4) {
    for (unused in seq_len(4 - length(series_ids))) {
      plot.new()
    }
  }
  par(mar = c(0, 0, 0, 0))
  plot.new()
  legend("center", horiz = TRUE, bty = "n",
         fill = c("#1b6da8", NA), border = c(NA, NA),
         lty = c(NA, 1), lwd = c(NA, 2), col = c(NA, "#c43c39"),
         legend = c("Observed", "Predicted"))
  layout(1)
  par(old_par)
} else {
  size_obs <- tanner$data$size_comp_observations
  series_ids <- sort(unique(size_obs$series))[1:4]
  old_par <- par(mfrow = c(2, 2), mar = c(3, 3, 2, 1), oma = c(0, 0, 2, 0))
  for (series in series_ids) {
    x <- size_obs[size_obs$series == series, ]
    x <- x[order(x$year), ]
    row <- x[nrow(x), ]
    bins <- grep("^bin_", names(row), value = TRUE)
    values <- as.numeric(row[1, bins])
    values <- values / sum(values)
    barplot(values, col = "#1b6da8", border = NA,
         xlab = "", ylab = "", main = paste(row$fleet, row$sex, row$catch_type, row$year),
         bty = "l")
    grid(col = "grey88")
  }
  mtext("Observed only: Tanner ADMB size-fit summary is not present", outer = TRUE, line = 0.4)
  par(old_par)
}
```
