---
title: "Tanner Crab GMACS Input Notes"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
execute:
  echo: true
  warning: false
  message: false
---

## Purpose

This document is the Tanner crab parallel to `bbrkc_rtmb_port_steps.qmd`.
The current Tanner work is at the input-orientation stage: the repository now
has a Tanner-specific reader for the available GMACS files in `examples/tanners`.

Unlike `examples/BBRKC`, the Tanner folder currently does not include
`Gmacsall.out` or `gmacs.rep`, so ADMB likelihood, penalty, prior, and
deterministic report parity cannot yet be audited.

## Files

The Tanner reader uses:

| File | Role |
|---|---|
| `examples/tanners/gmacs_26_22_03d5_aEffExp1.Run1.dat` | GMACS main run file. |
| `examples/tanners/TannerCrab_Data202509.EffExp1.dat` | Tanner data file. |
| `examples/tanners/TannerCrab_26_22_03d5_a.Run1.ctl` | Tanner control file. |
| `examples/tanners/TannerCrab_26_22_03d5_a.prj` | Tanner projection file. |
| `examples/tanners/gmacs_26_22_03d5_aEffExp1.par` | Fitted parameter table. |

Optional ADMB parity targets are discovered if they are added to the same
folder. Canonical names are `Gmacsall.out` and `gmacs.rep`; the reader also
checks common Tanner-specific variants such as
`gmacs_26_22_03d5_aEffExp1.Gmacsall.out` and
`gmacs_26_22_03d5_aEffExp1.rep`.

## Current Reader

The Tanner files use a different GMACS text layout from the BBRKC example.
`R/read_gmacs_tanner.R` therefore reads the Tanner file set directly instead of
forcing these inputs through BBRKC-specific parser assumptions.

```{r}
source("R/read_gmacs_bbrkc.R")
source("R/read_gmacs_tanner.R")

tanner <- read_tanner_inputs("examples/tanners")
validate_tanner_inputs(tanner)
```

```{r}
data.frame(
  report = names(tanner$admb_reference$files),
  path = unlist(tanner$admb_reference$files),
  available = file.exists(unlist(tanner$admb_reference$files)),
  row.names = NULL
)
```

## Input Summary

```{r}
dims <- tanner$data$dimensions
data.frame(
  quantity = names(dims),
  value = as.integer(dims),
  row.names = NULL
)
```

```{r}
data.frame(
  component = c("catch", "survey", "size composition", "growth", "maturity"),
  dataframes = c(
    tanner$data$catch$n,
    tanner$data$survey$n,
    tanner$data$size_comp$n,
    NA_integer_,
    NA_integer_
  ),
  rows = c(
    sum(tanner$data$catch$rows),
    sum(tanner$data$survey$rows),
    sum(tanner$data$size_comp$rows),
    tanner$data$growth$n,
    tanner$data$maturity$n
  )
)
```

## Parameter Summary

```{r}
data.frame(
  parameter_rows = nrow(tanner$parameters),
  active_estimated_parameters = sum(tanner$parameters$active),
  min_estimated_index = min(tanner$parameters$estimated_index, na.rm = TRUE),
  max_estimated_index = max(tanner$parameters$estimated_index, na.rm = TRUE)
)
```

```{r}
head(tanner$parameters[tanner$parameters$active, c(
  "parameter_index", "estimated_index", "value", "label"
)], 12)
```

## Current Status

The Tanner reader validates:

- model dimensions and size breaks;
- fishery and survey names;
- catch, survey, and size-composition dataframe counts;
- growth and maturity observation counts;
- fitted parameter table extraction from the `.par` file.

The next implementation step is to add Tanner ADMB report outputs
(`Gmacsall.out` and `gmacs.rep`, or their Tanner-equivalent names) so the same
style of objective and deterministic parity checks used for BBRKC can be built.
The reader now has the hooks to parse those files when present:

- `read_admb_likelihood_reference()` for `nloglike`, `nlogPenalty`, and
  `priorDensity`;
- `read_admb_catch_fit_summary()`, `read_admb_index_fit_summary()`, and
  `read_admb_size_fit_summary()` for observed/predicted fit summaries;
- `read_admb_selectivity_block()` for `slx_capture`, `slx_retaind`, and
  `slx_discard` from the report file.

## Index Diagnostics

The available Tanner folder does not include predicted index fits. The figure
below shows the observed index series that the future Tanner RTMB port should
match.

```{r}
#| label: fig-tanner-index-observations
#| fig-cap: "Observed Tanner crab survey index series."
#| code-fold: true
index_obs <- tanner$data$survey_observations
series_ids <- sort(unique(index_obs$series))
old_par <- par(mfrow = c(ceiling(length(series_ids) / 2), 2), mar = c(3, 3, 2, 1), oma = c(1, 1, 0, 0))
for (series in series_ids) {
  x <- index_obs[index_obs$series == series, ]
  plot(x$year, x$obs, type = "b", pch = 16, col = "#1b6da8",
       xlab = "", ylab = "", main = paste(x$fleet[1], x$sex[1], x$maturity[1]),
       bty = "l")
  grid(col = "grey88")
}
par(old_par)
```

## Size-Composition Diagnostics

The panels below show observed Tanner size compositions for recent rows from
the first four size-composition dataframes.

```{r}
#| label: fig-tanner-size-composition-observations
#| fig-cap: "Observed Tanner crab size compositions for selected recent rows."
#| code-fold: true
size_obs <- tanner$data$size_comp_observations
series_ids <- sort(unique(size_obs$series))[1:4]
old_par <- par(mfrow = c(2, 2), mar = c(3, 3, 2, 1), oma = c(1, 1, 0, 0))
for (series in series_ids) {
  x <- size_obs[size_obs$series == series, ]
  x <- x[order(x$year), ]
  row <- x[nrow(x), ]
  bins <- grep("^bin_", names(row), value = TRUE)
  values <- as.numeric(row[1, bins])
  values <- values / sum(values)
  barplot(values, col = "#1b6da8", border = NA,
       xlab = "", ylab = "", main = paste(row$fleet, row$sex, row$catch_type, row$year),
       bty = "l")
  grid(col = "grey88")
}
par(old_par)
```
